---
title: "Data Frames in R"
author: "Jai Keshav Sharma"
date: "2025-03-18"
output: github_document
---

## Data Frames

-   A data frame is R’s most natural way of presenting a data set with a collection of recorded observations for one or more variables.
-   **List vs Data Frame**: The most important distinction is that in a data frame (unlike a list), the members must all be vectors of equal length.
-   If you pass vectors of unequal length to data.frame, then R will attempt to recycle any shorter vectors to match the longest, throwing your data off and potentially allocating observations to the wrong variable.

### Creating a Data Frame

to create a data frame, **data.frame()** function is used. It takes vectors as input and combines them to form a data frame.

```{r}
mydata <- data.frame(person = c("John", "Doe", "Jane"),
                     age = c(23, 45, 34),
                     weight = c(70, 80, 65))
mydata
```

### Extraction

```{r}
mydata[2, 2]
```

```{r}
mydata[1:3, 3]
```

```{r accessing named variables}
mydata$age
```

### Size of a Data Frame

-   You can report the size of a data frame—the number of records and variables—just as you’ve seen for the dimensions of a matrix.

```{r}
dim(mydata)
```

```{r}
nrow(mydata)
```

```{r}
ncol(mydata)
```

### Character Vectors into Factors

-   R’s default behavior for character vectors passed to data.frame is to save it as character vector.

```{r}
str(mydata)
```

`str()`: It can display even the internal structure of large lists which are nested. It provides one-liner output for the basic R objects letting the user know about the object and its constituents.

-   If you want all strings to be converted to factors, you need to use the flag `stringsAsFactors` to TRUE.

```{r}
mydata <- data.frame(person = c("John", "Doe", "Jane"),
                     age = c(23, 45, 34),
                     weight = c(70, 80, 65),
                     stringsAsFactors = TRUE)
mydata
```

```{r}
str(mydata)
```

### Adding Data to Data Frame

1.  using `rbind()`: adding a row to the data frame

```{r}
newdata <- data.frame(person = "Sara", age = 41, weight = 60)
mydata <- rbind(mydata, newdata)
mydata
```

2.  using `cbind()` : adding a column to the data frame

```{r}
funny <- factor(x = c("High", "Low", "Medium", "Low"), levels = c("Low", "Medium", "High"))

mydata <- cbind(mydata, funny)
mydata
```

3.  Expanding the data frame by adding a new column:

```{r}
mydata$height <- c(170, 180, 160, 165)
mydata
```

### Extending data frame

```{r}
mydata$age.mon <- mydata$age * 12
mydata
```

### Logical Record Subsets

-   We can use logical flag vectors to subset data structures.
-   This is a particularly useful technique with data frames, where you’ll often want to examine a subset of entries that meet certain criteria.

```{r}
mydata[mydata$age > 30, ]
```

```{r}
mydata[mydata$age > 30 & mydata$weight < 70, ]
```

```{r}
mydata[mydata$person == "Doe", ]
```

```{r}
mydata[mydata$age > 30, c("person", "age", "weight")]
```

### Remove data from Data Frame

***Removing row***:

```{r}
mydata <- subset(mydata, person != "Doe")
mydata
```

\*we used the `subset()` function and specified the condition `person != Doe`. This removed the row with `person` equal to 3.

***Removing column***:

```{r}
mydata <- mydata[, -3]
mydata
```
