---
output: html_notebook
---

# Characters

```{r}
foo <- "this is a character string"
foo2 <- c("this is a character string", "hi")
length(foo)  ## not suitable for finding number of characters in a string
length(foo2)
```

### Using nchar() to find length of string

```{r}
nchar(foo)
nchar(foo2)

```

```{r}
temp <- "22.2"
temp
temp * 2  # Error in temp * 2 : non-numeric argument to binary operator
```

### String Comparison

```{r}
"hello" == "hello"
"Hello" == "hello"
"hello" == "apple"
```

#### Note: There is no concept of ASCII values in comparison of strings in R. Its follows a different criteria of comparing strings.

#### It order is : aA \< bB \< cC \< dD \< ... \< zZ

#### Also,  there is no such order for special characters is defined in R.

```{r}
"hello" > "apple"
"Hello" > "Apple"
"apple" > "Apple" 
"Apple" > "abble"

```

```{r}
"hello" < "hello " ## not checking length, doing character by character comparison
"!" < "@"
"@" < "#"  ## no order known for special charcters
"@" == "@"
```

### Concatenation

```{r echo=FALSE}
str1 <- cat("hola", "amigos")
str1
```

```{r}
str2 <- paste("hola", "amigos")
str2
```

```{r}
cat("hi\nall") ## new line is added

```

```{r}
paste("hi\nall") ## no escape sequence expanded because paste has to return its                           content to a variable
```

```{r}
"hi\nall" ## no escape sequence expanded as this notation can also be used to store in              variable
```

```{r}
class(cat("hi\nall")) ## NULL bcz cat returns nothing
```

```{r}
class(paste("hi\nall"))  ## "character"
```

```{r}
paste("hi", "how", "are", "you", "!", sep = "--")  
paste("hi", "how", "are", "you", "!", sep = "") 
paste("hi", "how", "are", "you", "!") ## default separator is " " (space)
```

### Escape Sequences in R

```{r}
/n
/t
/b
//
/"
```
